``Interestingness'' in Mathematical Theory Exploration

### 2019-05-22 PhDThesis/background_quickcheck.tex
Headingy things
Property checking rather than QuickCheck

\subsection{QuickCheck}
\label{sec:quickcheck}

@@

The intent of the function is to map an input $n$ to an output $n!$. The tests
check a few properties of the implementation, including the base case, that the
function is monotonically increasing, and a relationship between adjacent
outputs. However, these tests will \emph{not} expose a serious problem with the
implementation: it diverges on half of its possible inputs!

@@

Haskell offers a partial solution to this problem in the form of \emph{property
  checking}. Tools such as \qcheck{} separate tests into three components: a
\emph{property} to check, which unlike a unit test may contain \emph{free
  variables}; a source of values to instantiate these free variables; and a
stopping criterion. Here is how we might restate our unit tests as properties:

@@

Property checking is certainly an improvement over unit testing, but the problem
of tests being biased towards expected cases remains, since we are manually
specifying the properties to be checked.

We can reduce this bias further through the use of \emph{theory exploration}
tools, such as \qspec{} and \hspec{}. These programs \emph{discover} properties
of a ``theory'' (e.g. a library), through a combination of brute-force
enumeration, random testing and (in the case of \hspec{}) automated theorem
proving.

>>>>

\section{Property Checking}
\label{sec:propertychecking}

% TODO: Automated testing
% TODO: Agile
% TODO: Unit testing
% TODO:

@@

The intent of the function is to map an input $n$ to an output $n!$. The tests
check a few properties of the implementation, including the base case, that the
function is monotonically increasing, and a relationship between adjacent
outputs. However, these tests will \emph{not} expose a serious problem with the
implementation: it diverges on half of its possible inputs.

@@

Haskell offers a partial solution to this problem in the form of \emph{property
  checking}. Tools such as \qcheck{} separate tests into three components: a
\emph{property} to check, which unlike a unit test may contain \emph{free
  variables}; a source of values to instantiate these free variables; and a
stopping criterion.

\subsection{\qcheck{}}
\label{sec:quickcheck}

\qcheck{}~\cite{quickcheck} is the most widely used property checking library
for
Haskell~\footnote{According to downloads counted at
  \href{http://hackage.haskell.org/packages/browse}, and packages depending on
  it at \href{https://packdeps.haskellers.com/reverse/QuickCheck}, accessed on
  2019-05-21.}. Here is how we might restate our unit tests as \qcheck{}
properties:

@@

Whilst property checking generalises and improves upon unit testing, the problem
of tests being biased towards expected cases remains, since the properties to be
checked are still manually specified. Property checking can be complemented by
\emph{theory exploration} (TE), which avoids this bias by \emph{discovering}
such properties; through a combination of brute-force enumeration, random
testing and (in the case of \hspec{} and \hipster{}) automated theorem proving.
These TE systems make use of property checkers as part of their search process,
to try and disprove generated conjectures before output (to the user, or a
subsequent tool).
